%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%   BornAgain User Manual
%%
%%   homepage:   http://www.bornagainproject.org
%%
%%   copyright:  Forschungszentrum Jülich GmbH 2015
%%
%%   license:    Creative Commons CC-BY-SA
%%
%%   authors:    Scientific Computing Group at MLZ Garching
%%               C. Durniak, M. Ganeva, G. Pospelov, W. Van Herck, J. Wuttke
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\cleardoublepage
\ichapter{Preface}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\isection{About BornAgain}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

BornAgain is a software package
to simulate and fit
reflectometry, off-specular scattering,
and grazing-incidence small-angle scattering (GISAS)
of X-rays and neutrons.
It provides a generic framework
for modeling multilayer samples with smooth or
rough interfaces and with various types of embedded nanoparticles.
The name, \BornAgain,
alludes to the central role of the distorted-wave Born
approximation (DWBA) in the physical description of the
scattering process.
\index{Distorted-wave Born approximation}

BornAgain is maintained
by the Scientific Computing Group
of the J\"ulich Centre for Neutron Science (JCNS)
at Heinz Maier-Leibnitz Zentrum (MLZ) Garching, Germany.
It free and open source software.
The source code is released under the GNU General Public License (GPL, version 3 or higher),
the documentation under the Creative Commons license CC-BY-SA.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\isection{Citation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The canonical reference for BornAgain is the journal article
\begin{quote}
Gennady Pospelov, Walter Van Herck, Jan Burle, Juan M. Carmona Loaiza,
Céline Durniak, Jonathan M. Fisher, Marina Ganeva, Dmitry Yurov and
Joachim Wuttke:\\
BornAgain: software for simulating and fitting
grazing-incidence small-angle scattering\\
\href{https://doi.org/10.1107/S1600576719016789}{J. Appl. Cryst. 53, 262–276 (2020)}
\end{quote}
Use of the software version should additionally be documented by citing a specific version
\index{Citation}%
\begin{quote}
BornAgain --- Software for simulating and fitting
X-ray and neutron small-angle scattering at grazing incidence,
version $\langle$\texttt{version}$\rangle$ ($\langle$\texttt{release date}$\rangle$),\\
\url{http://www.bornagainproject.org}
\end{quote}
Citation of the present document is only necessary
when referring to specific information about form factors.
