\newpage
\chapter{Quick start} \SecLabel{QuickStart}

\section{Quick start on Unix Platforms}

This section shortly describes how to build and install \BornAgain\ 
from source and run the first simulation on Unix Platforms. 
Further details about the installation procedure are given in \SecRef{Installation}. \\

\noindent
{\bf Step I: $~$ install the third party software}
\begin{itemize}
\item compilers: clang  versions $\geq 3.1$ or GCC versions $\geq 4.2$
\item cmake ($\geq 2.8$)
\item boost library ($\geq 1.48$)
\item GNU scientific library ($\geq 1.15$)
\item fftw3 library ($\geq 3.3.1$)
\item \Python-2.7, python-devel, python-numpy-devel
%\item Eigen3 library ($\geq 3.1.0$), optional
%\item ROOT framework ($\geq 5.34.00$), optional
\end{itemize}
\vspace*{2mm}


\noindent
{\bf Step II: $~$ get the source} \newline
Download \BornAgain\ source tarball from \url{http://apps.jcns.fz-juelich.de/BornAgain}
or use the following git repository
\begin{lstlisting}[language=shell, style=commandline]
git clone git://apps.jcns.fz-juelich.de/BornAgain.git 
\end{lstlisting}

\vspace*{3mm}



\noindent
{\bf Step III: $~$ build the libraries and executable}
\begin{lstlisting}[language=shell, style=commandline]
mkdir <build_dir>; cd <build_dir>;
cmake -DCMAKE_INSTALL_PREFIX=<install_dir> <source_dir>
make
make check
make install
\end{lstlisting}
\vspace*{3mm}


\noindent
{\bf Step IV: $~$ run an example}
\begin{lstlisting}[language=shell, style=commandline]
python <install_dir>/share/BornAgain/Examples/python/simulation/ex001_CylindersAndPrisms/CylindersAndPrisms.py
\end{lstlisting}



\section{Quick start on Windows Platforms}

\noindent
{\bf Step I: $~$ install the third party software} \newline
The current version of \BornAgain\ requires \Code{Python, numpy, matplotlib} 
to be installed on the system. If you don't have them already installed,
you can use \Code{PythonXY} installer available 
at \url{https://code.google.com/p/pythonxy} which, with default installation options, contains at least these three packages.
%\BornAgain\ installation.
\vspace*{2mm}

\noindent
{\bf Step II: $~$ use BornAgain installation package } \newline
Windows installation package can be downloaded from \url{http://apps.jcns.fz-juelich.de/BornAgain}.
Double-click on it to start the installation process. Then follow the instructions.
\vspace*{2mm}

\noindent
{\bf Step III: $~$ run the example} \newline
Run an example located in \BornAgain\ installation directory:
\begin{lstlisting}[language=shell, style=commandline]
python C:/BornAgain-0.9.2/Examples/python/simulation/ex001_CylindersAndPrisms/CylindersAndPrisms.py
\end{lstlisting}


\section{Getting help}
Users of the software who encounter problems during the installation
of the framework or during the run of a simulation can use the web-based issue tracking system
at
\url{http://apps.jcns.fz-juelich.de/redmine/projects/bornagain/issues}
to report a bug. The same system can be used to request new features.
This system is open for all users in read mode, while 
submitting bug reports and feature requests are possible only after a simple registration
procedure.




%Requirements

%Hardware
%BornAgain is known to work on following platforms:
%Linux (x86, amd64)
%MacOS X (x86)

%Software
%GCC 4.1.2 or above   C/C++ compiler
%or
%clang
%gcc 4.1.2 or above, clang
